---
title: "ST 558 Project 2"
author: "Melanie Kahn & Bennett McAuley"
output:
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(httr)
library(jsonlite)
library(tidyverse)
```


## Overview

The purpose and overall goal of this vignette is provide instruction on how to contact an API using functions we created to query, parse, and return data in a well-structured format. These functions drive a basic exploratory analysis.

This vignette makes use of the beer data API by the [Open Brewery DB](https://www.openbrewerydb.org/faq) project.

## Requirements

The following packages should be installed and loaded to run the code successfully:

- `httr` - Tools for working with URLs and HTTP
- `jsonlite` - JSON parser and generator for R
- `tidyverse` - Collection of packages for data science using "tidy data"

## The Data

This section provides the context and structure of the data to get a better understanding of the contents of the API.

The brewery data from Open Brewery DB features information on breweries, cideries, brewpubs, and bottleshops around the world. Using the `GET` function from the `httr` package, a list of breweries is returned.

```{r}
library(httr)
breweries <- GET("https://api.openbrewerydb.org/breweries")
str(breweries, max.level = 1)
```

This information isn't very readable, so we will parse it using the `FromJSON` function from the `jsonlite` package, and represent the information as a tibble.

```{r}
brew_parsed <- fromJSON(rawToChar(breweries$content))
as_tibble(brew_parsed)
```

The variables present for each entry in this database are as follows:

- `id` - The unique ID of the brewery
- `name`
- `brewery_type` - The type of brewery; must be one of:
  - `micro` - Most craft breweries
  - `nano` - An extremely small brewery that typically only distributes locally
  - `regional` - Regional location of an expanded brewery
  - `brewpub` - A beer-focused restaurant or bar with a brewery on-premise
  - `planning` - A brewery in planning or not yet opened to the public
  - `contract` - A brewery that uses another brewery's equipment
  - `proprietor` - Similiar to contract brewing but refers more to a brewery incubator
  - `closed` - A location that has closed
- `street` - The street address of the brewery
- `address_2`
- `address_3`
- `city`
- `state`
- `county_province`
- `postal_code`
- `country`
- `longitude` - The vertical distance from an origin point
- `latitude` - The horizontal distance from an origin point
- `phone`
- `website_url`
- `updated_at`
- `created_at`

Notice that some of the variables provide useful information that can be applied to data analysis; some cannot (`website_url`, for instance). This will be addressed by the functions (_see below_) that will query the data and return tibbles that are analysis-ready.

[Back to Top](#top)

## Function Definitions

This section is dedicated to showcasing all of the functions go into contacting the API, querying data, and performing our basic exploratory analysis.

### Get_OB_DataFrame

The `Get_OB_DataFrame` function contacts the API, runs a query based on values provided by the user, parses the data from the query, and returns a tibble with relevant and well-formatted variables. It takes the following arguments:

- `size` - The number of breweries to return as observations (default is `10`)
- `search_by` - The category of filtering to be applied to the search. Valid options are:
  - `"city"`
  - `"state"`
  - `"country"`
  - `"type"` (_`brewery_type`, see above for values_)
  - `"name"`
  - `"dist"` (_must be denoted by `"<latitude>,<longitude>"`--e.g. `"45,37"`)
- `input` - The value or term to be searched for

_Note: In testing, it was discovered that the querying functionality is not case sensitive (i.e. `"raleigh"`, `"Raleigh"`, and `"RALEIGH"` will all return the same thing), so inputting the values in sentence case is not necessary. Inputting them in quotations on function call, however, is._

```{r API contact}
Get_OB_DataFrame <- function(size = 10, search_by, input) {
  if (search_by %in% c("city", "state", "country", "type", "name", "dist")) {
    query <- GET(paste0("https://api.openbrewerydb.org/breweries?by_", search_by, "=", input, "&per_page=", size))
  } else {
    stop("Invalid search category. Please use one of these options: 'city', 'state', 'country', 'type' 'name', or 'dist'.")
  }

  query_parse <- fromJSON(rawToChar(query$content))
  
  dt <- as_tibble(query_parse) %>%
    select(id, name, brewery_type, city, state, county_province, country, latitude, longitude) %>%
    mutate(brewery_type = as.factor(brewery_type), longitude = as.numeric(longitude), latitude = as.numeric(latitude)) %>%
    arrange(name)
  
  return(dt)
}
```

#### Usage Examples:
```{r df usage, eval = FALSE}
Get_OB_DataFrame(size = 50, search_by = "city", input = "Dublin") #returns 50 breweries in cities called 'Dublin'--not just the one in Ireland!
Get_OB_DataFrame(7, "state", "North Carolina") #returns 7 breweries in North Carolina
Get_OB_DataFrame(search_by = "country", input = "South Korea") #returns default number of breweries in South Korea
Get_OB_DataFrame(size = 20, "type", "nano") #return 20 nano breweries
Get_OB_DataFrame(search_by = "name", input = "wine") #returns default number of breweries with the term 'wine' contained within their names
Get_OB_DataFrame(size = 5, "dist", "41.9,12.4") #returns 5 closest breweries (that have non-NA values) to specified coordinates -- Lat 41.9, Long 12.4 (Rome, Italy)
```
[Back to Top](#top)

### Get_OB_Random

`Get_OB_Random` is functionally identical to `Get_OB_DataFrame`, but instead of the user specifying characteristics they want in the data, the observations are queried at random based on the following argument:

- `n` - The number of breweries to return as observations (default is `5`). According to the API documentation, the maximum is `50`.

_Note: In testing, it was verified that the randomness in the API is not really random; there is a seed associated with it as the queries yielded the same results per value of `n`, regardless of whether a seed was speicifed in the R environment._

```{r Random}
Get_OB_Random <- function(n = 5) {
  
  query <- GET(paste0("https://api.openbrewerydb.org/breweries/random?size=", n))

  query_parse <- fromJSON(rawToChar(query$content))
  
  dt <- as_tibble(query_parse) %>%
    select(id, name, brewery_type, city, state, county_province, country, latitude, longitude) %>%
    mutate(brewery_type = as.factor(brewery_type), longitude = as.numeric(longitude), latitude = as.numeric(latitude)) %>%
    arrange(name)
    
  
  return(dt)
}
```

#### Usage Examples
```{r random usage, eval = FALSE}
Get_OB_Random() #returns default number of randomly selected breweries
Get_OB_Random(20) #returns 20 randomly selected breweries
Get_OB_Random(75) #returns 50 randomly selected breweries; API will still execute the query, but caps the number of observations returned at 50

```
[Back to Top](#top)

## Exploratory Data Analysis

### Pulling the Data
_at least two calls of obtaining data function(s) (possibly combining into one)_

First, we will pull data from the API to create new tibbles using the `Get_OB_DataFrame()` function to contact, query, and parse the API data. `CA_Brew` is a new tibble with 50 breweries in the state of California; `IreEng_Brew` is a new tibble with 100 breweries, 50 in Ireland and 50 in England; and `USA_Brew` is a new tibble with 50 breweries located within the United States.

```{r}
CA_Brew <- Get_OB_DataFrame(size = 50, search_by = "state", input = "California")
CA_Brew

IRE_Brew <- Get_OB_DataFrame(size = 50, search_by = "country", input = "Ireland")
ENG_Brew <- Get_OB_DataFrame(size = 50, search_by = "country", input = "England")
IreEng_Brew <- rbind(IRE_Brew, ENG_Brew)
IreEng_Brew

USA_Brew <- Get_OB_DataFrame(size = 50, search_by = "country", input = "United States")
USA_Brew
```

### Manipulations

_create at least one new variable that is a function of other vars_

A new variable column named `regions` is added to the California breweries tibble `CA_Brew` that further categorizes the breweries according to their location in California. Historically, the state is separated into Northern and Southern California based on their geography. The Northern California region is mountainous and volcanic causing more dynamic weather, whereas Southern California is flat with gradual faults and arid deserts. The administrative line that connects San Luis Obispo, Kern, and San Bernardino counties creates the boundary between the two regions. The latitude measurement for this boundary line is 35.8°N. Therefore, to create this new variable, and brewery with a latitude less than or equal to 35.8 is located in the Southern region of California while anything greater than 35.8 is located in the Northern region of California.

```{r}
CA_Brew$regions <- ifelse(CA_Brew$latitude <= 35.8, "Southern", "Northern")
CA_Brew
```

A new variable column named `regions` is also added to the USA breweries tibble `USA_Brew` that further categorizes the breweries according to their location in the United States. Using the Census Bureau's [list of states by regions and division](https://www2.census.gov/geo/pdfs/maps-data/maps/reference/us_regdiv.pdf), a vector was created for each region and used to populate the `regions` variable by searching the `state` column for any state corresponding with it. Any row with a `state` not corresponding to a region will return `ERROR` in the new column.

```{r}
USA_Brew <- USA_Brew %>% 
  mutate(regions = if_else(state %in% c("Connecticut", "Maine", "Massachusetts", "New Hampshire", "Rhode Island", "Vermont", "New Jersey", "New York", "Pennsylvania"), "Northeast",
                           ifelse(state %in% c("Indiana", "Illinois", "Michigan", "Ohio", "Wisconsin", "Iowa", "Kansas", "Minnesota", "Missouri", "Nebraska", "North Dakota", "South Dakota"), "Midwest",
                                  if_else(state %in% c("Delaware", "District of Columbia", "Florida", "Geogria", "Maryland", "North Carolina", "South Carolina", "Virginia", "West Virginia", "Alabama", "Kentucky", "Mississippi", "Tennessee", "Arkansas", "Louisiana", "Oklahoma", "Texas"), "South",
                                          if_else(state %in% c("Arizona", "Colorado", "Idaho", "New Mexico", "Montana", "Utah", "Nevada", "Wyoming", "Alaska", "California", "Hawaii", "Oregon", "Washington"), "West", "ERROR")))))
USA_Brew
```
### Summaries

_some contingency tables, numerical summaries for some quantitative variables at each setting of some categorical variables (may not be inherently relevant to our data...)_

#### Categorical: Contingency Tables

The code chunk below creates a two-way contingency table showing the number of each type of brewery (from the `brewery_type` variable) per region (from the `regions` variable) for the 50 random breweries selected in California (`CA_Brew`) using the `table()` function. 

The first row of this contingency table indicates that of the 14 brewpubs observed in this data set, 5 are located in Northern California, while the other 9 are located in Southern California.

```{r}
table(CA_Brew$brewery_type, CA_Brew$regions)
```

The code chunk below creates a one-way contingency table showing the number of each type of brewery (from the `brewery_type` variable) in the data set with 50 random breweries selected in the United States (`USA_Brew`) using the `table()` function. 

This contingency table indicates that there are 9 brewpubs, 2 closed, 1 contract, 7 large, 30 micro, and 1 proprietor brewery in the USA data.

```{r}
table(USA_Brew$brewery_type)
```

The code chunk below creates a two-way contingency table showing the number of each type of brewery (from the `brewery_type` variable) per region (from the `regions` variable) for the 50 random breweries selected in the United States (`USA_Brew`) using the `table()` function. 

The first row of this contingency table indicates that of the 9 brewpubs observed in this data set, 1 is located in the Midwest, 2 are located in the Northeast, 2 are located in the South, and 4 are located in the West.

```{r}
table(USA_Brew$brewery_type, USA_Brew$regions)
```

The code chunk below creates a two-way contingency table showing the number of each type of brewery (from the `brewery_type` variable) per `country` for the 100 random breweries selected in Ireland and England (`IreEng_Brew`) using the `table()` function. 

This contingency table indicates that micro breweries are by far the most popular in both countries (32 in England and 36 in Ireland), taprooms are more popular in England (9 in England and none in Ireland), and regional breweries are more popular in Ireland (none in England and 4 in Ireland).

```{r}
table(IreEng_Brew$brewery_type, IreEng_Brew$country)
```

#### Quantitative: Numerical Summaries

Running the code chunk below creates a summary statistics table showing the average (`mean()`) and median (`median()`) for the latitude and longitude values in the United States breweries according to `state` rounded to the nearest hundredth. 

The second row indicates that, according to the 50 random breweries selected from the United States to make up the `USA_Brew` data set, breweries in California have average coordinates of (35.68°N, 119.63°W) and median coordinates of (35.86°N, 119.35°W).

```{r}
USA_Brew2 <- USA_Brew %>%
  group_by(state) %>%
  summarize(Avg_Lat = round(mean(latitude), digits = 2),
            Avg_Long = round(mean(longitude), digits = 2),
            Median_Lat = round(median(latitude), digits = 2),
            Median_Long = round(median(longitude), digits = 2))
USA_Brew2
```

### Plots

_at least 5 plots with nice labels and titles; one bar, one histogram, one box, and one scatter_

#### Bar Plot

Running the code chunk below creates a side-by-side bar plot to visualize the relationship between the brewery types (`brewery_type`) and location (`country`) for the observations in the Irish and English breweries data set (`IreEng_Brew`). The `stat` argument within `geom_bar` is set to `"count"`, `brewery_type` is set to the x-axis, and the `fill` is set to the `country` variable, telling R to count and plot the number each type of brewery per country. The `brewery_type` and `country` variables were titled accordingly on the x-axis and in the legend, the `count` variable has been capitalized, and the plot itself was titled `"Bar Plot of Breweries in Ireland and England"`.

As mentioned above in the discussion of this bar plot's corresponding two-way contingency table, the resulting bar graph again shows that most breweries in Ireland and England are microbreweries, indicating the craft beer is most popular in both countries.

```{r}
g <- ggplot(data = IreEng_Brew, aes(x = brewery_type, fill = country))
g + geom_bar(stat = "count", position = "dodge") +
  labs(title = "Bar Plot of Breweries in Ireland and England", x = "Type of Brewery", y = "Count") +
  scale_fill_discrete(name = "Country")
```

#### Histogram

Running the code chunk below creates a histogram to visualize the distribution of latitude values for observations in the USA breweries tibble (`latitude` in `USA_Brew`). Since there are 11 observations in the data set which have missing values for their coordinates, these NA values are dropped from the data before plotting (`na.rm = TRUE`). The width of the bins (`binwidth`) is set to 1 for better visualization of the data, while the colors of the bins and their outline are set to yellow and brown respectively. The `count` and `latitude` variables have been capitalized, and the plot itself was titled `"Histogram of Latitude Values for USA Breweries"`.

The resulting histogram shows that most breweries in the US are located between 39°N and 44°N, which is mostly in the northern United States.

```{r}
g2 <- ggplot(USA_Brew, aes(x = latitude))
g2 + geom_histogram(binwidth = 1, na.rm = TRUE, color = "brown", fill = "yellow") +
  labs(title = "Histogram of Latitude Values for USA Breweries", x = "Latitude", y = "Count")
```

